\bfseries Convolution-based nonlinear functions on tensor fields and tensor images

### 2023-03-20 paper/main.tex
Updates from Overleaf

% Copyright 2022 the authors

% To-Do list:
% -----------
% - What are our GOALS!!?
%   -- one option: this formulation can express all equivariant differential operators discretized to a cubical d-torus. 
%   -- another option: can we produce a characterization of all equivariant functions (maybe based on invariants or these filters) that don't require the use of non-linearities.
% - Understand what functions can and cannot be expressed on this model 
% - Should we be representing images in fourier space? Ask Leslie!
% - Understand the literature
%   -- Discretized vector calculus and differential geometry.
%   -- Group-equivariant CNNs; what's been done?
%   -- Understand how this formulation compares with steerable CNNs
%   -- What is known about what functions CNNs can represent? all linear translation equivariant functions can be written as convvolutions if the filters are as large as the image (simple observation)
% what about non-linear? is there a polynomial basis here? 
%  -- Define "local" functions, is there a characterization of functions that can be written as convolutions with small filters?
% - There is a one-indexing and zero-indexing issue. Should we be consistent?
% - Some definitions begin with "given" some with "let". Should we try to be more consistent?
% add zeroth order monomials (constants, very few)
% --- we don't need to be in the torus and we don't need to be square (or cubical)

%-- BEN: compute Molien's formula for this (translations semidirect D_4) action
% 1/|G|\sum_{g\in G} det(I-\phi(g)t)^{-1}

% SOLEDAD: audit the levi civita contraction
%GIVE THE METHOD A NAME AND RUN THE NAME THROUGH THE DOCUMENT
%reformat the monomial figures to look like equations
%visualize the group $B_d$ at d=2
%make an introduction figure showing examples of geometric filters

% - HOGG: Change the aspect ratios of the figures so they can be put into talk slides :)

%pixel positions in images are 1-0-tensors and they transform as 1-0-tensors

%change the notation from k-p to k_p
%should we use the upper index notation??
%t

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[letterpaper]{geometry} % because Overleaf is weird.
\PassOptionsToPackage{hyphens}{url}
\usepackage[hidelinks]{hyperref}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amsthm}

@@

% theorem, definition and so on
\theoremstyle{plain}
\newtheorem{definition}{Definition}
\newtheorem{conjecture}{Conjecture}
\newtheorem{theorem}{Theorem}

% math macros
% HOGG SAY: I DON'T KNOW HOW TO USE AMSMATH \choose AND I'M ON A PLANE. FEEL FREE TO FIX THIS.
\renewcommand{\choose}[2]{\begin{pmatrix}{#1}\\{#2}\end{pmatrix}}

% text macros
\newcommand{\sectionname}{Section}
\newcommand{\secref}[1]{\sectionname~\ref{#1}}
\newcommand{\figref}[1]{\figurename~\ref{#1}}

% typesetting adjustments
\makeatletter
\renewcommand\section{\@startsection {section}{1}{\z@}%
  {-3.25ex \@plus -1ex \@minus -.2ex}%
  {1.5ex \@plus .2ex}%
  {\raggedright\normalfont\large\bfseries}}% key thing is: \raggedright!
\makeatother
\setlength{\textwidth}{5.00in}
\setlength{\oddsidemargin}{3.25in}
\addtolength{\oddsidemargin}{-0.5\textwidth}
\setlength{\textheight}{9.40in}
\setlength{\topmargin}{-0.50in}
\setlength{\footskip}{0.25in}
\pagestyle{myheadings}
\markboth{foo}{Hogg, Blum-Smith, \& Villar / Geometric filters for geometric images}
\newcommand{\figurerule}{\rule[1ex]{\textwidth}{0.2pt}}
\linespread{1.08}
\frenchspacing\sloppy\sloppypar\raggedbottom

\title{\bfseries%
Convolution-based nonlinear functions on tensor fields and tensor images}
\author{David W. Hogg, Ben Blum-Smith, and Soledad Villar}
\date{}

@@

\paragraph{Abstract:}
Convolutional neural networks are the main tools for machine learning with images at the present day.
These methods effectively assume that the input images are scalar images (an intensity or brightness in each pixel, possibly in three or more channels), and they enforce a kind of locality and translation-equivariance.
In natural-science domains, image-like data sets can have a vector (velocity, say), or tensor (polarization, say), or pseudovector (magnetic field, say), or any combination in each pixel.
We provide a construction for general convolution-based functions of images or lattices
containing scalars, vectors, and tensors, producing scalar, vector, and tensor outputs.
Our formulation, which combines a geometric generalization of convolution with products, tensor index contractions, contractions with Levi-Civita symbols, and tensor index permutations, can universally approximate any suitably local analytic scalar, vector, or tensor functions on scalar, vector, or tensor images.
The framework permits, with a very simple adjustment, restriction to function spaces that are exactly equivariant to translations, discrete rotations, and reflections, but the framework will be useful even in contexts in which the functions are not expected to be equivariant.

@@

Contemporary natural science teems with data sets that are images or lattices or grids of geometric objects.
These might be observations of intensities (scalars), magnetic fields (pseudovectors), or polarizations (2-tensors) on a surface or in a volume.
Or they might be the inputs or outputs of a simulation, in which the initial conditions or fields are specified on a regular grid.
For examples, see \figref{fig:examples} and the caption thereto.
Any lattice or grid of vectors or tensors can be seen as a generalization of the concept of an image, in which the intensity in each pixel is replaced with a geometric object (scalar, vector, pseudovector, tensor) in each pixel.
These objects are \emph{geometric} in the sense that they are defined in terms of their transformation properties under geometric operators such as rotation, translation, and reflection.
Thus there is a need for machine learning methods designed for \emph{geometric images}---lattices or grids of scalars, vectors, and tensors.
There are already countless applications of machine learning in contexts in which the input data are geometric images, including examples in essentially all natural-science disciplines (HOGG CITE THINGS).
\begin{hoggfigure}
  \begin{center}
    \begin{minipage}[b]{2.5in}\noindent%
      \includegraphics[width=\textwidth]{jpegPIA18916.jpeg}\\
      \textsl{(a)} temperature and polarization
    \end{minipage}
    \begin{minipage}[b]{2in}\noindent%
      \includegraphics[width=\textwidth]{key_figures_197.pdf}\\
      \textsl{(b)} salinity and current
    \end{minipage}\\[2ex]
    \begin{minipage}[b]{2.5in}\noindent%
      \includegraphics[width=\textwidth]{applsci-09-02108-g001.png}\\
      \textsl{(c)} temperature
    \end{minipage}
    \begin{minipage}[b]{1.8in}\noindent%
      \includegraphics[width=\textwidth]{The-distribution-of-the-modified-potential-vorticity-at-the-850-K-potential-temperature.png}\\
      \textsl{(d)} vorticity
    \end{minipage}\\[2ex]
    \begin{minipage}[b]{1.9in}\noindent%
      \includegraphics[width=\textwidth]{FbAjqkkWAAANqFc.jpeg}\\
      \textsl{(e)} intensity and polarization
    \end{minipage}
    ~~\begin{minipage}[b]{2.5in}\noindent%
      \includegraphics[width=\textwidth]{41524.pdf}\\[1ex]
      \textsl{(f)} stress tensor
    \end{minipage}
  \end{center}
  \caption{Examples of geometric images in the natural sciences.
  \textsl{(a)}~A visualization of a temperature map and a polarization map from the ESA \textsl{Planck} Mission \cite{planck}.
  The color map shows a temperature field (a scalar or 0-0-tensor) on the sphere, and the whiskers show the principal eigenvector direction of a 2-0-tensor field in two dimensions.
  In detail the underlying data are represented on a pixel grid (healpixel \cite{healpix}) on the sphere.
  \textsl{(b)}~Some other stuff from \url{https://climatedataguide.ucar.edu/climate-data/godas-ncep-global-ocean-data-assimilation-system} HOGG EXPLAIN AND CITE.
  \textsl{(c)}~Some other stuff from \url{https://www.mdpi.com/2076-3417/9/10/2108/htm} HOGG EXPLAIN AND CITE.
  \textsl{(d)}~From \url{https://www.researchgate.net/figure/The-distribution-of-the-modified-potential-vorticity-at-the-850-K-potential-temperature_fig3_26635879}.
  \textsl{(e)}~Ask Jordy Davelaar (and thank him).
  \textsl{(f)}~From \url{https://www.nature.com/articles/s41524-019-0234-8}.
  \label{fig:examples}}
\end{hoggfigure}

%For a concrete example close to one of our hearts (DWH's), contemporary cosmological n-body simulations comprise the most computationally expensive part of most contemporary cosmological projects, missions, and experiments (CITE THINGS).
%For this reason, accurate, fast emulators might be of enormous economic and scientific value.
%There are now quite a few successful projects implementing such emulators (CITE THINGS).
%Importantly, the initial conditions for these simulations are tiny displacements (vectors) and velocities (vectors) relative to a cubical lattice on a 3-torus.
%That is, the initial conditions can be described as a 3-dimensional image of vectors on a regular grid.
%For another concrete example, astronomical imaging data taken of the interstellar medium can be represented as rectangular lattices of measurements of temperature (scalar), optical depth (scalar), and induced polarization of starlight (a 2-tensor).
%Here the data can be described as a 2-dimensional image of scalars and tensors on a regular grid.

At the present day, the go-to tools for machine learning with images are convolutional neural networks (CNNs; CITE) and their many descendents (for example, CITE).
These methods are designed to work on one- or few-channel images in which, at the bottom (early) layers of the network, the operations involve image convolutions with learned filters followed by the application of pointwise nonlinearities.
In typical contexts, the channels of multi-channel input images will be something like the red, green, and blue (or cyan, magenta, yellow, and black) channels of a color image; these can be combined arbitrarily in the layers of the CNN.
When these CNN-based tools are applied to lattices of vectors (say), typically the components of the vectors are treated just as channels of the input image (for an example, see CITE) and then everything proceeds as with multi-channel color images.
This is not good!

@@

The fundamental observation inspiring this work is that when the components of vectors and tensors are acted on with arbitrary functions, the geometric structure of these objects is destroyed.
There are strict rules, dating back to the invention of differential geometry (CITE RICCI++), about how geometric objects can be combined to produce new geometric objects, consistent with coordinate freedom and transformation rules.
In previous work (CITE VILLAR++, YAO++) we have capitalized on these geometric rules to develop modified machine-learning methods that are restricted to exactly obey group-theoretic equivariances in physics contexts.
Here we use these rules to create a comprehensive set of tools that are universally approximating for functions that take geometric images as input and produce geometric labels (or geometric-image labels) as output.

We are motivated in this work to help solve problems in the natural sciences, where geometric images abound.
However, we conjecture that these tools are probably very useful even for standard machine-learning image-recognition and image-regression tasks.
After all, even standard images are measurements of intensity (a scalar) at a regular grid of points on a two-dimensional surface.
If the labels we seek are also scalars, it will behoove us to obey the rules of geometry.

These rules are roughly as follows:
A $k$-tensor object (tensor of order $k$) in $d$ dimensions has $k$ indices, each of which can take a value from 1 to $d$; that is, the $k$-tensor is an element of $(\mathbb R^d)^{\otimes k}$.
It can be contracted to a $(k-2)$-tensor object by identifying a pair of indices and summing over them.
A $k$-tensor and a $k'$-tensor can be multiplied (outer product) to make a $(k+k')$-tensor object, and then contractions can bring down the order.
$1$-tensor objects are called vectors and $0$-tensor objects are called scalars.
There are also odd-parity versions of all these (pseudovectors, pseudoscalars, and pseudotensors), and parity-changing contractions using the Levi-Civita symbol, so in what follows we will define $k$-$p$-tensors that have $k$ indices and a parity $p\in [0, 1]$.
Two objects can only be added or subtracted if they have the same order $k$ and parity $p$.
This will all be made precise below.
But in short, these rules define objects that can be given transformation rules under rotation and reflection such that functions made of these operations are coordinate free, or equivariant to any change of coordinate system (including curvilinear coordinate transformations).

The symmetries that suggest these rules are continuous symmetries.
But of course images are usually---and for our purposes---discrete grids of values.
This suggests that in addition to the continuous symmetries respected by the tensor objects in the image pixels there will be discrete symmetries for each geometric image taken as a whole.
We will define these discrete symmetry groups and use them to define a useful kind of group equivariance for functions of geometric images.
This equivariance, it turns out, is very easy to enforce, even for nonlinear functions of geometric images, provided that we compose our nonlinear functions from simple geometric operations, including our geometric generalization of convolution.
When we enforce this equivariance, the convolution filters that appear look very much like the differential operators that appear in discretizations of vector calculus, which is perhaps not surprising.

\paragraph{Related work:}

\paragraph{Our contribution:}

\section{Geometric objects and geometric images}\label{sec:geometric}

In this \sectionname{} we define the geometric objects and geometric images that we use to generalize classical images in scientific contexts.
The main point is that the channels of geometric images---which will be like the components of vectors and tensors---are not independent.
There is an underlying action by the orthogonal group that rules what are the allowed objects and transformations in this space.

We first fix $d$, the dimension of the space (typically 2 or 3 but it could be arbitrary). The geometric objects are vectors and tensors. 
The orthogonal group $O(d)$ can act on $v \in \mathbb R^d$ in the following ways:
\begin{align} \label{eq.action}
    g\cdot v = \det(M(g))^p M(g) v 
\end{align}
where $g\in O(d)$, $M(g)\in \mathbb R^{d\times d}$ is the standard matrix representation of $g$, and $p\in\{0,1\}$ is the parity of the action. If $p=0$ we obtain the standard $O(d)$ action on $\mathbb R^d$ \emph{vectors}. If $p=1$ we obtain what in physics is commonly known as the action of $O(d)$ on \emph{pseudovectors}. In this context the objects are defined by the actions that they carry.

\begin{definition}[$k$-$p$-tensors]\label{def.tensors}
We say that $v\in \mathbb R^d$ is a 1-$p$-tensor if $O(d)$ acts on $v$ via the action \eqref{eq.action}. 
If $v_i$ is a 1-$p_i$-tensor then $T:=v_{1}\otimes\ldots \otimes v_k$ is a rank-1 $k$-$p$-tensor, where $p=\sum_{i=1}^k p_i \mod 2$ and the action of $O(d)$ is defined as
\begin{align}
    g\cdot (v_{1}\otimes\ldots \otimes v_k) = (g\cdot v_1)\otimes \ldots \otimes (g\cdot v_k)
\end{align}
Higher rank $k$-$p$-tensors are defined as linear combinations of rank-1  $k$-$p$-tensors where the action of $O(d)$ is extended linearly.
\end{definition}
Typically we would make the parity $p$ be a bit (either 0 for even parity or 1 for odd parity) but in principle it can be any integer.

We note that in physics the 1-0-tensors are known as \emph{vectors}, the 1-1-tensors are known as \emph{pseudovectors}, the 0-0-tensors are known as \emph{scalars}, 0-1-tensors are known as \emph{pseudoscalars}, the $k$-1-tensors with $k\geq 2$ are known as \emph{pseudotensors}, and finally the $k$-0-tensors with $k\geq 2$ are the things that are commonly known as \emph{tensors}.
In general, any $k$-$p$-tensor $a$ can be written as a sum of outer products of order-$1$ ($k=1$) tensors (vectors and pseudovectors), where each term in the sum is an outer product of $k$ order-$1$ tensors and the parity $p$ is the sum of the parities of the input order-$1$ tensors (mod 2, perhaps).

\begin{definition}[summation notation]
In this notation, tensor products are written in component form, and repeated indices are summed over.
In this notation, the product of two 2-tensors (represented as two $d\times d$ matrices $A$ and $B$) is written as
\begin{equation}
    [A\, B]_{i,j} = [A]_{i,k}\,[B]_{k,j} ~,
\end{equation}
where $[A]_{i,k}$ (for example) is the $i,k$ element of matrix $A$, the repeated index $k$ is implicitly summed over the integers from 1 to $d$.
This notation works for tensor expressions of any order, provided that every index appears either exactly once (so it isn't summed over; in the above expression $i, j$ appear exactly once) or exactly twice (so it is summed over; in the above expression $k$ appears exactly twice). 
\end{definition}

One consequence of the above definitions is that 
$T\in (\mathbb R^d)^{\otimes k}$ is a $k$-$p$-tensor if and only if (in summation notation)
\begin{align}
    [g\cdot T]_{i_1,\ldots, i_k} = \det(M(g))^p\, [T]_{j_1,\ldots,j_k}\,[M(g)]_{i_1,j_1}\cdots[M(g)]_{i_k,j_k}
\end{align} for all $g\in O(d)$, where $[T]_{i_1, \ldots ,i_k} \in \mathbb R$ is a component of $T$, $[M(g)]_{i,j}\in\mathbb R$ is the $i,j$ element of the matrix representation of $g$, and all the $i_m$ and $j_m$ are indices in the range $1,\ldots,d$.
For example, a 2-0-tensor is defined by its transformation property
$[g\cdot T]_{i,j} = [T]_{k,\ell}\,[M(g)]_{k,i}\,[M(g)]_{\ell,j}$,
which, in normal matrix notation, is written as
$g\cdot T = M(g)\,T\,M(g)^\top$.

We consider an image $A$ in with $N$ equally spaced pixels in each dimension ($N^d$ pixels total). Each pixel contains a $k$-$p$-tensor ($k$ and $p$ are the same for each pixel). Sometimes we will consider the image to be in a $d$-torus. 
Let $\mathcal T_{d,k,p}$ the set of $k$-$p$-tensors in $\mathbb R^d$. We define the geometric images as follows.
\begin{definition}[geometric image]
A geometric image is a function $A:[N]^d \to \mathcal T_{d,k,p}$, where $[N]=\{0,1,\ldots, N-1\}$. We will also consider $k$-tensor images on the $d$-torus, where $[N]^d$ is given the algebraic structure of $(\mathbb Z / N\mathbb Z)^d$.
\end{definition}

\section{Generalized convolutions on geometric images}\label{sec:convolution}

>>>>

% Copyright 2022, 2023 the authors

% To-Do list:
% -----------

% - What are our GOALS!!? is it clearly stated in the abstract and introduction?
% - Understand what functions can and cannot be expressed on this model

% MARIA TERESA:
% - Understand the following very specific things in the literature:
%   -- Discretized vector calculus and differential geometry.
%   -- Group-equivariant CNNs; what's been done?
%   -- Understand how this formulation compares with steerable CNNs
%   -- What is known about what functions CNNs can represent? all linear translation equivariant functions can be written as convvolutions if the filters are as large as the image (simple observation) what about non-linear?  
%  -- comment on the fact that Cohen and Kondor showed that in homogeneous spaces all linear equivariant maps are convolutions, say in what sense we generalize the result, by combining it with equivariant functions on tensors and vectors and scalars can be written in terms of scalars. 
%  -- Define "local" functions, is there a characterization of functions that can be written as convolutions with small filters?
% -- in the introduction we should talk about the geometric principle

% SOLEDAD:
% - Some definitions begin with "given" some with "let". Should we try to be more consistent?
% Define the constants like the dirac delta and the levi civita,
% - COMMENT: In the same way that we can take nonlinear functions of scalars, we can take ODD nonlinear functions of pseudoscalars (with no bias term permitted!).
% - write our contributions (there is a placeholder in the text).

% WILSON:
% - GIVE THE METHOD A NAME AND RUN THE NAME THROUGH THE DOCUMENT
% - Make a new figure that shows what happends when the filters are applied to a geometric image. Make it look like an equation or a rebus.
% - visualize the group $B_d$ at d=2
% - comment that \bar i is a \tensor{1}{+}
% - look at the different definitions of the action of g and see that they are consistent
% - add explicit comments after definitions, where appropriate. Use the "remark" environment. Look at Hogg's remarks for examples.
% - add a description of the numerical example problem
% - describe the possible architectures and a diagram

% - HOGG:
% - comment on alias alibi in the style of Weyl, both in the Introduction, and at the definition of the group action on a tensor.
% - Should we make the contravariant / covariant distinction here, or at least comment on it? Probably yes. should we use the upper index notation?? We should at least explain why people do that. Maybe we should.
% - Equation (1) is the ugliest thing in the Universe at present.

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[letterpaper]{geometry} 
\PassOptionsToPackage{hyphens}{url}
\usepackage[hidelinks]{hyperref}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{verbatim}
\usepackage{physics}

@@

% theorem, definition and so on
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{conjecture}{Conjecture}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem*{remark}{Remark}

% math macros
\newcommand{\tensorname}[2]{{#1}_{(#2)}}
\newcommand{\tensor}[2]{$\tensorname{#1}{#2}$-tensor}
% HOGG SAY: I DON'T KNOW HOW TO USE AMSMATH \choose AND I'M ON A PLANE. FEEL FREE TO FIX THIS.
\renewcommand{\choose}[2]{\begin{pmatrix}{#1}\\{#2}\end{pmatrix}}

% text macros
\newcommand{\sectionname}{Section}
\newcommand{\secref}[1]{\sectionname~\ref{#1}}
\newcommand{\figref}[1]{\figurename~\ref{#1}}
\newcommand{\wgg}[1]{\textcolor{purple}{#1}}

% typesetting adjustments
\makeatletter
\renewcommand\section{\@startsection {section}{1}{\z@}%
  {-3.25ex \@plus -1ex \@minus -.2ex}%
  {1.5ex \@plus .2ex}%
  {\raggedright\normalfont\large\bfseries}}% key thing is: \raggedright!
\makeatother
\setlength{\textwidth}{5.00in}
\setlength{\oddsidemargin}{3.25in}
\addtolength{\oddsidemargin}{-0.5\textwidth}
\setlength{\textheight}{9.40in}
\setlength{\topmargin}{-0.50in}
\setlength{\footskip}{0.25in}
\pagestyle{myheadings}
\markboth{foo}{Gregory, Hogg, Blum-Smith, Arias, \& Villar / Geometric Image CNNs}
\newcommand{\figurerule}{\rule[1ex]{\textwidth}{0.2pt}}
\linespread{1.08}
\frenchspacing\sloppy\sloppypar\raggedbottom

\title{\bfseries%
Convolution-based nonlinear functions on tensor fields and tensor images}
\author{%
Wilson~Gregory,
David~W.~Hogg,
Ben~Blum-Smith,
\\
Maria~Teresa~Arias,
Soledad~Villar
}
\date{}

@@

\paragraph{Abstract:}
Convolutional neural networks are the main tools for machine learning with images at the present day.
These methods effectively assume that the input images are scalar images (an intensity or brightness in each pixel, possibly in three or more channels), and they enforce a kind of locality and translation-equivariance.
In natural-science domains, image-like data sets can have a vector (velocity, say), or tensor (polarization, say), or pseudovector (magnetic field, say), or any combination in each pixel.
Here we provide a construction for general convolution-based functions of images or lattices
containing scalars, vectors, and tensors, producing scalar, vector, and tensor outputs.
Our formulation, which combines a geometric generalization of convolution with products, tensor index contractions, contractions with Levi-Civita symbols, and tensor index permutations, can universally approximate any suitably local analytic scalar, vector, or tensor functions on scalar, vector, or tensor images.
The framework permits, with a very simple adjustment, restriction to function spaces that are exactly equivariant to translations, discrete rotations, and reflections, but the framework will be useful even in contexts in which the functions are not expected to be equivariant.

@@

Contemporary natural science is replete with data sets that are images, lattices, or grids of geometric objects.
These might be observations of intensities (scalars), magnetic fields (pseudovectors), or polarizations (2-tensors) on a surface or in a volume.
They might be the inputs or outputs of a simulation where the initial conditions or fields are specified on a regular grid; see \figref{fig:examples} for some examples.
Any lattice of vectors or tensors can be seen as a generalization of the concept of an image in which the intensity in each pixel is replaced with a geometric object (scalar, vector, pseudovector, tensor).
These objects are \emph{geometric} in the sense that they are defined in terms of their transformation properties under geometric operators such as rotation, translation, and reflection.
Thus there is a need for machine learning methods designed for \emph{geometric images}---lattices or grids of scalars, vectors, and tensors.
There are already countless applications of machine learning in contexts in which the input data are geometric images, including examples in essentially all natural-science disciplines.
\begin{hoggfigure}
  \begin{center}
    \begin{minipage}[b]{2.5in}\noindent%
      \includegraphics[width=\textwidth]{jpegPIA18916.jpeg}\\
      \textsl{(a)} temperature and polarization
    \end{minipage}
    \begin{minipage}[b]{2in}\noindent%
      \includegraphics[width=\textwidth]{key_figures_197.pdf}\\
      \textsl{(b)} salinity and current
    \end{minipage}\\[2ex]
    \begin{minipage}[b]{2.5in}\noindent%
      \includegraphics[width=\textwidth]{applsci-09-02108-g001.png}\\
      \textsl{(c)} temperature
    \end{minipage}
    \begin{minipage}[b]{1.8in}\noindent%
      \includegraphics[width=\textwidth]{The-distribution-of-the-modified-potential-vorticity-at-the-850-K-potential-temperature.png}\\
      \textsl{(d)} vorticity
    \end{minipage}\\[2ex]
    \begin{minipage}[b]{1.9in}\noindent%
      \includegraphics[width=\textwidth]{FbAjqkkWAAANqFc.jpeg}\\
      \textsl{(e)} intensity and polarization
    \end{minipage}
    ~~\begin{minipage}[b]{2.5in}\noindent%
      \includegraphics[width=\textwidth]{41524.pdf}\\[1ex]
      \textsl{(f)} stress tensor
    \end{minipage}
  \end{center}
  \caption{Examples of geometric images in the natural sciences.
  \textsl{(a)}~A visualization of a temperature map and a polarization map from the ESA \textsl{Planck} Mission \cite{planck2015}.
  The color map shows a temperature field (a scalar or \tensor{0}{+}) on the sphere, and the whiskers show the principal eigenvector direction of a \tensor{2}{+} field in two dimensions.
  In detail the underlying data are represented on a pixel grid (healpixel \cite{healpix}) on the sky (a 2-sphere).
  \textsl{(b)}~Two-dimensional maps of ocean current (shown with arrows; a vector or \tensor{1}{+} field) and ocean salinity (shown with color; a scalar or \tensor{0}{+} field) at a depth of 5\,m \cite{climatedataguide}.
  \textsl{(c)}~A three-dimensional map of temperature (a scalar or \tensor{0}{+} field) based on sensors distributed throughout the volume of a granary \cite{granary}.
  \textsl{(d)}~A two-dimensional map of potential vorticity (a pseudoscalar or \tensor{0}{-} field) in the Earth's atmosphere, measured for the purposes of predicting storms \cite{potentialvorticity}.
  \textsl{(e)}~Two-dimensional maps on the sky of intensity $I$ and the three independent components $Q, U, V$ of the electromagnetic polarization \tensor{2}{+}, from a simulation of a jet outflow from an accreting black hole [Davelaar et al, in preparation].
  \textsl{(f)}~Components of the three-dimensional stress tensor (a \tensor{2}{+} field) in a diamond anvil cell, which is used to study the behavior of samples at exceedingly high pressures \cite{diamond}.
  \label{fig:examples}}
\end{hoggfigure}

At the present day, the go-to tools for machine learning with images are convolutional neural networks (CNNs; \cite{lecun1989backpropagation}) and their many descendents (for example, CITE).
These methods are designed to work on one- or few-channel images in which the early layers of the network involve image convolutions with learned filters followed by the application of pointwise nonlinearities.
In typical contexts, the channels of multi-channel input images will be something like the red, green, and blue (or cyan, magenta, yellow, and black) channels of a color image; these can be combined arbitrarily in the layers of the CNN.
When these CNN-based tools are applied to lattices of vectors for example, typically the components of the vectors are treated just as channels of the input image (for an example, see CITE) and then everything proceeds as with multi-channel color images.
This is not good! %WILSON: I don't love this. HOGG says: You are the first author. Be bold; edit.

@@

The fundamental observation inspiring this work is that when the components of vectors and tensors are acted on with arbitrary functions, the geometric structure of these objects is destroyed.
There are strict rules, dating back to the early days of differential geometry \cite{ricci}, about how geometric objects can be combined to produce new geometric objects, consistent with coordinate freedom and transformation rules (these rules constitute a theme of \cite{mcp}).
In previous work \cite{villar2021scalars, villar2022dimensionless, yao2021simple} we have capitalized on these geometric rules to develop modified machine-learning methods that are restricted to exactly obey group-theoretic equivariances in physics contexts.
Here we use these rules to create a comprehensive set of tools that are universally approximating for functions that take geometric images as input and produce geometric images as output.

We are motivated in this work to help solve problems in the natural sciences, where geometric images abound.
However, we conjecture that these tools are probably very useful even for standard machine-learning image-recognition and image-regression tasks.
After all, even standard images are measurements of intensity (a scalar) at a regular grid of points on a two-dimensional surface.
If the labels we seek are also scalars, it behooves us to obey the rules of geometry.

These rules are roughly as follows:
A $k$-tensor object (tensor of order $k$) in $d$ dimensions has $k$ indices, each of which can take a value from 1 to $d$; that is, the $k$-tensor is an element of $(\mathbb R^d)^{\otimes k}$.
It can be contracted to a $(k-2)$-tensor object by identifying a pair of indices and summing over them.
A $k$-tensor and a $k'$-tensor can be multiplied (outer product) to make a $(k+k')$-tensor object, and then contractions can bring down the order.
$1$-tensor objects are called vectors and $0$-tensor objects are called scalars.
There are also odd-parity versions of all these (pseudovectors, pseudoscalars, and pseudotensors), and parity-changing contractions using the Levi-Civita symbol, so in what follows we will define \tensor{k}{p}s that have $k$ indices and a parity $p\in\{-1,+1\}$, sometimes denoted ``$-$'' and ``$+$'' below.
Two objects can only be added or subtracted if they have the same order $k$ and parity $p$.
These rules define objects that can be given transformation rules under rotation and reflection such that functions made of these operations are coordinate free, or equivariant to any change of coordinate system, including curvilinear coordinate transformations.

The symmetries that suggest these rules are continuous symmetries.
But of course images are usually---and for our purposes---discrete grids of values.
This suggests that in addition to the continuous symmetries respected by the tensor objects in the image pixels there will be discrete symmetries for each geometric image taken as a whole.
We will define these discrete symmetry groups and use them to define a useful kind of group equivariance for functions of geometric images.
This equivariance, it turns out, is very easy to enforce, even for nonlinear functions of geometric images, provided that we compose our nonlinear functions from simple geometric operations, including our geometric generalization of convolution.
When we enforce this equivariance, the convolution filters that appear look very much like the differential operators that appear in discretizations of vector calculus.

\paragraph{Our contribution:}
SOLEDAD: SUMMARY HERE. 
To see a comparison of our work with other methods already studied, see \sectionname~\ref{sec:related}
For a discussion of prior work on group equivariant machine learning, see \sectionname~\ref{sec:related2}.

\section{Non-linear functions of geometric images}

We define the geometric objects and geometric images that we use to generalize classical images in scientific contexts in \sectionname{} \ref{sec:geometric} and \sectionname{} \ref{sec:convolution}, and we define the geometric functions on them in \sectionname{} \ref{sec:nonlinear}.
The main point is that the channels of geometric images---which will be like the components of vectors and tensors---are not independent.
There is an underlying action by the orthogonal group that rules what are the allowed objects and transformations in this space that we describe in \sectionname{}~\ref{sec:equivariant}.

\subsection{Geometric objects}\label{sec:geometric}

We first fix $d$, the dimension of the space (typically 2 or 3 but it could be arbitrary).
The geometric objects are vectors and tensors. 
The orthogonal group $O(d)$ is the space of isometries of $\mathbb R^d$ that fix the origin. 
%It is represented by matrices $Q\in \mathbb R^{d\times d}$ such that $Q^\top Q = I$. 
It acts on vectors and pseudovectors $v \in \mathbb R^d$ in the following way:
\begin{align} \label{eq.action}
    g\cdot v = \det(M(g))^{\frac{1-p}{2}}\,M(g)\,v
\end{align}
where $g\in O(d)$, $M(g)\in \mathbb R^{d\times d}$ is the standard matrix representation of $g$ (i.e. $M(g)^\top M(g) = I$) and $p\in\{-1,+1\}$ is the parity of $v$.
If $p=+1$ we obtain the standard $O(d)$ action on $\mathbb R^d$ \emph{vectors}.
If $p=-1$ we obtain the $O(d)$ action on what in physics are known as \emph{pseudovectors}.
In this context the objects are defined by the actions that they carry.

\begin{definition}[\tensor{k}{p}s]\label{def.tensors}
We say that $v\in \mathbb R^d$ is a \tensor{1}{p} if $O(d)$ acts on $v$ via the action \eqref{eq.action}. 
If $v_i$ is a \tensor{1}{p_i} then $T:=v_{1}\otimes\ldots \otimes v_k$ is a rank-1 \tensor{k}{p}, where $p=\prod_{i=1}^k p_i$ and the action of $O(d)$ is defined as
\begin{align}
    g\cdot (v_{1}\otimes\ldots \otimes v_k) = (g\cdot v_1)\otimes \ldots \otimes (g\cdot v_k)
\end{align}
Higher rank \tensor{k}{p}s are defined as linear combinations of rank-1 \tensor{k}{p}s where the action of $O(d)$ is extended linearly. The set of \tensor{k}{p}s in $d$ dimensions is denoted $\mathcal{T}_{d,k,p}$.
\end{definition}

\begin{remark}[terminology and notation]
The parity $p$ is a signed bit, either $+1$ for even parity or $-1$ for odd parity.
Note that we call $k$ the \emph{order} of the \tensor{k}{p}, rather than the ``rank'' of the tensor as some other sources do. Using rank for this property is inaccurate because there can be \tensor{2}{+}s of rank 1 (like those we use in Definition~\ref{def.tensors}).
\end{remark}

\begin{remark}[odd-parity objects]
It is possible to replace any odd-parity ($p=-1$) object with an even-parity object with the same physical content by multiplying the odd-parity object by the Levi-Civita tensor (see Definition~\ref{def:levicivita}).
When a \tensor{k}{p} is multiplied (outer product) by a Levi-Civita, the result is a \tensor{(k+d)}{-p}.
Thus, it is possible to work without odd-parity objects at all, in principle.
However, since odd-parity objects are of importance in physics and engineering (see \figurename~\ref{fig:examples}), we retain them.
\end{remark}

In physics the \tensor{1}{+}s (such as velocities) are known as \emph{vectors},
the \tensor{1}{-}s (such as angular momenta) are known as \emph{pseudovectors},
the \tensor{0}{+}s (such as rest masses) are known as \emph{scalars},
the \tensor{0}{-}s (such as surface vorticities) are known as \emph{pseudoscalars},
the \tensor{k}{-}s with $k\geq 2$ are known as \emph{pseudotensors},
and finally the \tensor{k}{+}s with $k\geq 2$ are the things that are commonly known as \emph{tensors}.
In general, any \tensor{k}{p} can be written as a sum of outer products of order-$1$ ($k=1$) tensors (vectors and pseudovectors), where each term in the sum is an outer product of $k$ order-$1$ tensors and the parity $p$ is the product of the parities of the input order-$1$ tensors.

%CONSTANT TENSORS
We consider two special tensors that will be important for the definition of our models, the Kronecker delta and the Levi-Civita.

\begin{definition}[Kronecker delta] The Kronecker delta ($\delta_{ij}$) is a \tensor{2}{+} represented by the identity matrix, namely the object with two indices $ij$ such that it has the value $+1$ when the two indices have the same value ($i=j$), and $0$ otherwise.
\end{definition}

\begin{definition}[Levi-Civita symbol]\label{def:levicivita}
The Levi-Civita symbol in $d\geq 2$ is a \tensor{d}{-} $\epsilon_{i_1,\ldots, i_d}$ such that if the $d$ indices are not repeated and in an even-permutation order the value is $+1$ and if the $d$ indices are not repeated and in an odd-permutation order the value is $-1$, and it has the value $0$ in all other cases.
\end{definition}
